Overfitting example using hurricane frequencies

### 2021-12-12 text/21/hurricane-frequencies.Rmd
Updated hurricane-frequencies.Rmd

```{r notes}
# Note: this code can produce a reasonable PowerPoint
# presentation. The system has an annoying tendency
# to put the graphs and the speaker notes on separate
# slides.
```

@@

This [dataset on hurricane frequency][dasl1] is found at the [Data and Story Library website][dasl2].

@@

```{r poly1}
add_trend <- function(point_plot, df, degree) {
  x_grid <- data.frame(x=seq(1850, 2010, by=0.1))
  lm(y~poly(x, degree), data=df) %>%
    augment(newdata=x_grid) -> yhat
  yhat_2005 <- filter(yhat, x==2005)
  prediction <<- yhat_2005$.fitted
  point_plot +
    geom_line(data=yhat, aes(x, .fitted), color="red") +
    geom_label(
      data=yhat_2005, 
      aes(x, .fitted, label=round(.fitted, 1)))
}
add_trend(point_plot, df, 1)
```

@@

@@

@@

This code reads data directly from the website and prints out the first eight rows.

@@

This code is for printing the rest of the data.

@@

@@

It took a bit of work to put everything in a function, but now you can produce a quadratic rather than a linear trend with just a single line of code.

@@

Here's code for fitting the remaining trend lines.

>>>>

```{r notes, echo=FALSE}
# Note: this code can produce a reasonable PowerPoint
# presentation. The system has an annoying tendency
# to put the graphs and the speaker notes on separate
# slides.
```

@@

@@

```{r poly1}
add_trend <- function(point_plot, df, degree) {
  x_grid <- data.frame(x=seq(1855, 2005, by=0.1))
  lm(y~poly(x, degree), data=df) %>%
    augment(newdata=x_grid) -> yhat
  yhat_2005 <- filter(yhat, x==2005)
  prediction <<- yhat_2005$.fitted
  point_plot +
    geom_line(data=yhat, aes(x, .fitted)) +
    geom_label(
      data=yhat_2005, 
      aes(x, .fitted, label=round(.fitted, 1)))
}
add_trend(point_plot, df, 1)
```

@@

### Code (load libraries)

@@

### Code (read and print)

@@

@@

This code reads data directly from the website and prints out the data in two pieces.

@@

### Code (plot data points)

@@

@@

It took a bit of work to put everything in a function, but now you can produce a quadratic or higher level trend rather than a linear trend with just a single line of code.
